---
title: "ESM 244 Week 1 Lab"
author: "Anna Talken"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(dplyr)
```

### Read in the SF trees data
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```


### Basic Wrangling Reminders

Refresh some skills for data wrangling and summary statistics using functions in the 'dplyr' package

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.
```{r}
top_5_status <- sf_trees %>%
  count(legal_status) %>% 
  drop_na(legal_status) %>% #remove N/A row
  rename(tree_count = n) %>% # new name on the left, old name on the right
  relocate(tree_count) %>% # moved tree count to be the first column
  slice_max(tree_count, n = 5) #selecting top 5 groups for tree counts
```

Make a graph of top 5 observations

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y= tree_count)) +
    geom_col() + #fct_reorder reorders the bars to be in order from low to high tree count based on legal status
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() +
  theme_minimal()
```

### A few more data wrangling refreshers

Only want to keep observations (rows) for blackwood acacia

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>%  #filter to only keep rows where within the species variable there is the phrase 'Blackwood Acacia'
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x=longitude, y = latitude)) +
  geom_point() 
```

### tidyr:: separate() and unite() functions

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example of tidyr:: unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id: legal_status, sep = "_cool!_")
```


